---
output: 
  rmarkdown::github_document:
    html_preview: true
    fig_width: 4
    fig_height: 4
    toc: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
# Installation

Install ggplot2 and devtools if you don't have them already.

```{r eval = FALSE}
install.packages("ggplot2")
install.packages("devtools")
```

Install ggfittext from github.

```{r eval = FALSE}
library(devtools)
install_github('wilkox/ggfittext')
```

# Fitting text to a box

Sometimes you want to draw some text in ggplot2 so that it doesn't spill
outside a bounding box. For example, this doesn't look very good:

```{r doesnt_fit}
library(ggfittext)
flyers <- data.frame(
  vehicle = c("cessna", "jumbo jet", "space shuttle", "dyson sphere"),
  xmin = c(45, 40, 25, 10),
  xmax = c(55, 60, 75, 90),
  ymin = c(5, 35, 65, 95),
  ymax = c(25, 55, 85, 115)
)
ggplot(flyers) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax)) +
  geom_text(aes(label = vehicle, x = (xmin + xmax) / 2, y = (ymin + ymax) / 2))
```

ggfittext provides a geom called `geom_fit_text` that will shrink text (when
needed) to fit a bounding box.

```{r geom_fit_text}
ggplot(flyers, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                   label = vehicle)) +
  geom_rect() +
  geom_fit_text()
```

You can define the box with ‘xmin’ and ‘xmax’ aesthetics, or alternatively with
‘x’ and ‘width’ (width is given in millimetres). Likewise, you can use either
‘ymin’ and ‘ymax’ or ‘y’ and ‘height’. The ‘width’ and ‘height’ aesthetics
can be useful when drawing on a discrete axis.

You can specify where in the bounding box to place the text with `place`, and a
minimum size for the text with `min.size`. (Any text that would need to be
smaller than `min.size` to fit the box will be hidden.)

```{r geom_fit_text_2}
ggplot(flyers, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                   label = vehicle)) +
  geom_rect() +
  geom_fit_text(place = "topleft", min.size = 8)
```

Text can be placed in any corner or at the midpoint of any side (‘topleft’,
‘top’, ‘topright’, ‘right’…), as well as the default ‘centre’.

# Growing text

With the `grow = TRUE` argument, text will be grown as well as shrunk to fit the
box:

```{r geom_fit_text_3}
ggplot(flyers, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                   label = vehicle)) +
  geom_rect() +
  geom_fit_text(grow = TRUE)
```

# Reflowing text

The `reflow = TRUE` argument causes text to be reflowed (wrapped) as needed to fit
the bounding box. Reflowing is preferred to shrinking; that is, if the text can
be made to fit by reflowing it without shrinking it, it will be reflowed only.

```{r reflow}
poem <- data.frame(
  text = rep(
    "Whose text this is I think I know.\nHe would prefer it to reflow",
    3
  ),
  xmin = rep(10, 3),
  xmax = rep(90, 3),
  ymin = rep(10, 3),
  ymax = rep(90, 3),
  fit = c("geom_text", "no reflow", "reflow")
)
ggplot(poem, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                 label = text)) +
  geom_rect() +
  geom_text(
    data = subset(poem, fit == "geom_text"),
    aes(x = (xmin + xmax) / 2, y = (ymin + ymax) / 2)
  ) +
  geom_fit_text(data = subset(poem, fit == "no reflow"), min.size = 0) +
  geom_fit_text(data = subset(poem, fit == "reflow"), reflow = TRUE, min.size = 0) +
  lims(x = c(0, 100), y = c(0, 100)) +
  labs(x = "", y = "") +
  facet_wrap(~ fit)
```

Note that existing line breaks in the text are respected.

If both `reflow = TRUE` and `grow = TRUE` are passed, the text will be reflowed to
a form that best matches the aspect ratio of the bounding box, then grown to fit
the box.

```{r reflow_and_grow}
film <- data.frame(
  text = rep("duck soup", 3),
  xmin = rep(30, 3),
  xmax = rep(70, 3),
  ymin = rep(0, 3),
  ymax = rep(100, 3),
  fit = c("geom_text", "grow, no reflow", "grow and reflow")
)
ggplot(film, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                 label = text)) +
  geom_rect() +
  geom_text(
    data = subset(film, fit == "geom_text"),
    aes(x = (xmin + xmax) / 2, y = (ymin + ymax) / 2)
  ) +
  geom_fit_text(data = subset(film, fit == "grow, no reflow"), grow = TRUE) +
  geom_fit_text(
    data = subset(film, fit == "grow and reflow"),
    grow = TRUE,
    reflow = TRUE
  ) +
  lims(x = c(0, 100), y = c(0, 100)) +
  labs(x = "", y = "") +
  facet_wrap(~ fit, ncol = 1)
```
