---
output: 
  github_document:
    html_preview: true
    fig_width: 4
    fig_height: 4
---

```{r echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```
<!-- badges: start -->
[![R-CMD-check](https://github.com/wilkox/ggfittext/workflows/R-CMD-check/badge.svg)](https://github.com/wilkox/ggfittext/actions)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ggfittext)](https://cran.r-project.org/package=ggfittext)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

# ggfittext

ggfittext is a ggplot2 extension for fitting text into boxes.

```{r hero, echo = FALSE, message = FALSE, fig.width = 8, fig.height = 4}
yeats <- data.frame(
  xmin = c(0, 4, 6, 4, 4, 5, 5.5, 5,   5,    5.25, 5.25),
  xmax = c(4, 8, 8, 6, 5, 6, 6,   5.5, 5.25, 5.5,  5.5),
  ymin = c(0, 4, 0, 0, 2, 3, 2,   2,   2.5,  2.75, 2.5),
  ymax = c(8, 8, 4, 2, 4, 4, 3,   2.5, 3,    3,    2.75),
  label = c("<span style='color:red'>T</span><span style='color:orange'>u</span><span style='color:yellow'>r</span><span style='color:green'>n</span><span style='color:blue'>i</span><span style='color:purple'>n</span><span style='color:violet'>g</span>",
            "**and**", "*turning*", "<span style='color:blue'>in</span>", 
            "the<sup>*</sup>", "widening", "gyre", "the",
            "falcon", "cannot", "hear"),
  angle = c(0, 315, 270, 225, 180, 135, 90, 45, 0, 315, 270)
)
library(ggfittext)
library(ggplot2)
library(patchwork)
p1 <- ggplot(yeats, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, label =
                  label, angle = angle)) +
  geom_rect(fill = "grey", colour = "black") +
  geom_fit_text(grow = TRUE, min.size = 0, rich = TRUE)

polar_d <- data.frame(
  word = c("The" , "falcon" , "cannot" , "hear" , "the" , "falconer") ,
  xmin = c( 0.0  , 0.2      , 0.35      , 0.6    , 0.65   , 0.8)        ,
  xmax = c( 0.2  , 0.35     , 0.6      , 0.65    , 0.8   , 1)        ,
  ymin = rep(0, 6),
  ymax = c( 0.95 , 0.85      , 0.75     , 0.65    , 0.45  , 0.35)
)
p2 <- ggplot(polar_d, aes(label = word, xmin = xmin, xmax = xmax, ymin =
                         ymin, ymax = ymax)) +
  geom_rect(aes(xmin = 0, xmax = 1, ymin = 0, ymax = 1)) +
  geom_rect(fill = "lightblue", colour = "black") +
  coord_polar() +
  geom_fit_text(
    grow = TRUE,
    padding.x = grid::unit(1, "mm"),
    padding.y = grid::unit(1, "mm"),
    place = "top",
    min.size = 0
  )
p1 + p2
```

## Installation

Install the release version of ggfittext from CRAN:

```{r eval = FALSE}
install.packages("ggfittext")
```

If you want the development version, install it from GitHub:

```{r eval = FALSE}
devtools::install_github("wilkox/ggfittext")
```

## Fitting text inside a box

Sometimes you want to draw some text in a ggplot2 plot so that it fits inside a
defined area. You can do this by manually fiddling with the font size, but this
is tedious and un-reproducible. ggfittext provides a geom called
`geom_fit_text()` that automatically resizes text to fit inside a box.  It
works like this:

```{r, fig.width = 4, fig.height = 4}
ggplot(animals, aes(x = type, y = flies, label = animal)) +
  geom_tile(fill = "white", colour = "black") +
  geom_fit_text()
```

As with `geom_text()`, the position of the text is set by the `x` and `y`
aesthetics. `geom_fit_text()` tries to infer the width and height of the box in
which the text is allowed to fit, and shrinks down any text that is too big.

## Reflowing text

Another way to make the text fit in the box is by reflowing it; that is,
wrapping it over multiple lines. With the `reflow = TRUE` argument,
`geom_fit_text()` will reflow the text before (if still necessary) shrinking
it:

```{r, fig.width = 4, fig.height = 4}
ggplot(animals, aes(x = type, y = flies, label = animal)) +
  geom_tile(fill = "white", colour = "black") +
  geom_fit_text(reflow = TRUE)
```

## Growing text

If you want the text to be as large as possible, the argument `grow = TRUE`
will increase the text size to the maximum that will fit in the box. This works
well in conjunction with `reflow`:

```{r, fig.width = 4, fig.height = 4}
ggplot(animals, aes(x = type, y = flies, label = animal)) +
  geom_tile(fill = "white", colour = "black") +
  geom_fit_text(reflow = TRUE, grow = TRUE)
```

## Placing text

By default, text is placed in the centre of the box. However, you can place it
in a corner or on a side of the box with the `place` argument, which takes
values like "top", "topright", "bottomleft" and so on:

```{r, fig.width = 4, fig.height = 4}
ggplot(animals, aes(x = type, y = flies, label = animal)) +
  geom_tile(fill = "white", colour = "black") +
  geom_fit_text(place = "topleft", reflow = TRUE)
```

## Bar plots

ggfittext also provides a geom `geom_bar_text()` for labelling bars in bar
plots:

```{r, fig.width = 4, fig.height = 4}
ggplot(altitudes, aes(x = craft, y = altitude, label = altitude)) +
  geom_col() +
  geom_bar_text()
```

`geom_bar_text()` works with stacked bar plots:

```{r, fig.width = 4, fig.height = 4}
ggplot(beverages, aes(x = beverage, y = proportion, label = ingredient,
                    fill = ingredient)) +
  geom_col(position = "stack") +
  geom_bar_text(position = "stack", reflow = TRUE)
```

And it works with dodged bar plots, and with flipped bar plots:

```{r, fig.width = 4, fig.height = 4}
ggplot(beverages, aes(x = beverage, y = proportion, label = ingredient,
                    fill = ingredient)) +
  geom_col(position = "dodge") +
  geom_bar_text(position = "dodge", grow = TRUE, reflow = TRUE, 
                place = "left") +
  coord_flip()
```

### Experimental feature: rich text

With the `rich = TRUE` argument, `geom_fit_text()` and `geom_bar_text()` both
support a limited subset of Markdown and HTML markup for text (rendered with
[gridtext](https://wilkelab.org/gridtext/)).

```{r, fig.width = 4, fig.height = 4}
ggplot(animals_rich, aes(x = type, y = flies, label = animal)) +
  geom_tile(fill = "white", colour = "black") +
  geom_fit_text(reflow = TRUE, grow = TRUE, rich = TRUE)
```

Rich text cannot be drawn in polar coordinates. Please note that this feature
is liable to change, and is subject to upstream changes to gridtext.

## Specifying the box limits

If you want to manually set the limits of the box (instead of having them
inferred from `x` and `y`), you can use `xmin` & `xmax` and/or `ymin` & `ymax`:

```{r, fig.width = 4, fig.height = 4}
ggplot(presidential, aes(ymin = start, ymax = end, x = party, label = name)) +
  geom_fit_text(grow = TRUE) +
  geom_errorbar(alpha = 0.5)
```

Alternatively, you can set the width and/or height with the `width` and/or
`height` arguments, which should be `grid::unit()` objects. The horizontal
and/or vertical centre of the box will be defined by `x` and/or `y`.

## Experimental feature: text in polar coordinates

Text can be drawn in polar coordinates with `geom_fit_text()` simply by adding
`coord_polar()` to the plot. This feature is experimental and any bug reports
are very welcome.

```{r, fig.width = 4, fig.height = 4}
p <- ggplot(gold, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, 
                 fill = linenumber, label = line)) +
  coord_polar() +
  geom_rect() +
  scale_fill_gradient(low = "#fee391", high = "#238443")

p + geom_fit_text(min.size = 0, grow = TRUE)
```

When text is drawn in polar coordinates, the `flip = TRUE` argument can be used
to flip upside-down text the 'right way up' to ease readability:

```{r, fig.width=4, fig.height=4}
p + geom_fit_text(min.size = 0, grow = TRUE, flip = TRUE)
```

## Other useful arguments

All arguments to `geom_fit_text()` can also be used with `geom_bar_text()`.

- **`contrast`** can be used to automatically invert the colour of the text so
  it contrasts against a background `fill`:

```{r, fig.width = 4, fig.height = 4}
ggplot(animals, aes(x = type, y = flies, fill = mass, label = animal)) +
  geom_tile() +
  geom_fit_text(reflow = TRUE, grow = TRUE, contrast = TRUE)
```

- **`padding.x`** and **`padding.y`** can be used to set the padding between
  the text and the edge of the box. By default this is 1 mm. These values must
  be given as `grid::unit()` objects.
- **`min.size`** sets the minimum font size in points, by default 4 pt. Text
  smaller than this will be hidden (see also `outside`).
- **`outside`** is `FALSE` by default for `geom_fit_text()`. If `TRUE`, text
  that is placed at "top", "bottom", "left" or "right" and must be shrunk
  smaller than `min.size` to fit in the box will be flipped to the outside of
  the box (if it fits there). This is mostly useful for drawing text inside
  bars in a bar plot.
- **`hjust`** and **`vjust`** set the horizontal and vertical justification of
  the text, scaled between 0 (left/bottom) and 1 (right/top). These are both
  0.5 by default.
- **`formatter`** allows you to provide a function that will be applied to the
  text before it is drawn. This is mostly useful in contexts where variables
  may be interpolated, such as when using
  [gganimate](https://gganimate.com/).
- **`fullheight`** is automatically set depending on place, but can be
  overridden with this option. This is used to determine the bounding box
  around the text. If `FALSE`, the bounding box includes the x-height of the
  text and ascenders, but not any descenders. If TRUE, it extends from the top
  of the ascenders to the bottom of the descenders. This is mostly useful in
  situations where you want to ensure the baseline of text is consistent
  between labels (`fullheight = FALSE`), or when you want to avoid descenders
  spilling out of the bounding box (`fullheight = TRUE`).

```{r echo = FALSE, fig.width = 4, fig.height = 0.5}
ggplot(data.frame(label = "fullheight = TRUE", xmin = 0.25, xmax = 0.75, ymin = 0.25, ymax = 0.75), 
       aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, label = label)) +
  geom_rect() +
  geom_fit_text(grow = TRUE, fullheight = TRUE, family = "mono") +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), panel.background = element_blank(), plot.title = element_text(hjust = 0.5))
```

```{r echo = FALSE, fig.width = 4, fig.height = 0.5}
ggplot(data.frame(label = "fullheight = FALSE", xmin = 0.25, xmax = 0.75, ymin = 0.25, ymax = 0.75), 
       aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, label = label)) +
  geom_rect() +
  geom_fit_text(grow = TRUE, fullheight = FALSE, family = "mono") +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), panel.background = element_blank(), plot.title = element_text(hjust = 0.5))
```
