---
title: "ggfittext"
author: "David Wilkins"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# `geom_shrink_text`

Sometimes you want to draw some text in ggplot2 so that it doesn't spill outside
a bounding box. For example:

```{r}
library(ggfittext)
flyers
ggplot(flyers) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = class)) +
  geom_text(aes(label = vehicle, x = x, y = y))
```

ggfittext provides two new ggplot2 geoms that help with fitting text to a box.
To shrink text down so it doesn't overflow a bounding box, use
`geom_shrink_text`:

```{r}
ggplot(flyers, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, label =
                   vehicle, fill = class)) +
  geom_rect() +
  geom_shrink_text()
```

Note that, unlike with `geom_text`, you don't need to provide explicit ‘x’ and
‘y’ aesthetics to position the text label. Instead, you supply the position of
the bounding box with ‘xmin’, ‘xmax’, ‘ymin’, ‘ymax’ and position the text
within the box with ‘place’.

You can specify where in the bounding box to place the text, and hide text that
would be shrunk below a minimum size:

```{r}
ggplot(flyers, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, label =
                   vehicle, fill = class)) +
  geom_rect() +
  geom_shrink_text(place = "topleft", min.size = 8)
```

The obvious set of placements (‘topleft’, ‘top’, ‘topright’...) are supported,
as well as the default ‘centre’. ‘vjust’ and ‘hjust’ are set automatically.

# `geom_fill_text`

`geom_fill_text` will expand the text label to fix the bounding box. The text
aspect ratio will be maintained, so usually the size of the text will only be
maximised along one axis. As with `geom_shrink_text`, you specify the position
of the boundary box and the placement of the text label within the box.
`geom_fill_text` only support three places: ‘top’, ‘middle’ (the default) and
‘bottom’.

```{r}
ggplot(flyers, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, label =
                   vehicle, fill = class)) +
  geom_rect() +
  geom_fill_text()
```
