---
title: "ggfittext"
author: "David Wilkins"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Sometimes you want to draw some text in ggplot2 so that it doesn't spill outside
a bounding box. For example:

```{r}
library(ggfittext)
flyers
ggplot(flyers) +
  geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = class)) +
  geom_text(aes(label = vehicle, x = x, y = y))
```

ggfittext provides a new geom, `geom_fit_text`, that will shrink text when
needed to fit a bounding box:

```{r}
ggplot(flyers, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, label =
                   vehicle, fill = class)) +
  geom_rect() +
  geom_fit_text()
```

You can define the x position and dimension of the box with ‘xmin’ and ‘xmax’,
or alternatively with ‘x’ and ‘width’ (width is given in millimetres). Likewise,
the y dimension and position can be defined with ‘ymin’ and ‘ymax’ or with ‘y’
and ‘height’. This can be useful when drawing on a discrete axis.

You can specify where in the bounding box to place the text, and hide text that
would be shrunk below a minimum size:

```{r}
ggplot(flyers, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, label =
                   vehicle, fill = class)) +
  geom_rect() +
  geom_fit_text(place = "topleft", min.size = 8)
```

The obvious set of placements (‘topleft’, ‘top’, ‘topright’...) are supported,
as well as the default ‘centre’. ‘vjust’ and ‘hjust’ are set automatically.

With the `grow = T` argument, text will be grown as well as shrunk to fit the
box:

```{r}
ggplot(flyers, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, label =
                   vehicle, fill = class)) +
  geom_rect() +
  geom_fit_text(grow = T)
```
