---
title: "Introduction to 'ggfittext'"
author: "David Wilkins"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ggfittext}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Fitting text inside a box

Sometimes you want to draw some text in a ggplot2 plot so that it fits inside a
defined area. For example, you might want to label tiles in a heat map without
letting the labels spill over into other tiles; or you might want to constrain
some point labels to imaginary boxes so they don't get too big. It's possible
to achieve this by manually fiddling with the text size, but this is both
tedious and un-reproducible.

ggfittext provides a special geom called `geom_fit_text()` that automates
fitting text inside a box. It works more or less like `ggplot2::geom_text()`,
but provides some additional aesthetics and options that let you specify the box
in which the text is to fit and how to make it fit.

```{r doesnt_fit}
library(ggplot2)
library(ggfittext)

flyers <- data.frame(
  vehicle = rep(c("kite", "jumbo jet", "space shuttle", "dyson sphere"), 2),
  xmin = rep(c(52.5, 45, 30, 0), 2),
  xmax = rep(c(67.5, 75, 90, 120), 2),
  ymin = rep(c(0, 15, 35, 60), 2),
  ymax = rep(c(10, 30, 55, 85), 2),
  geom = factor(c(rep("geom_text", 4), rep("geom_fit_text", 4)),
                levels = c("geom_text", "geom_fit_text"))
)

ggplot(flyers, aes(label = vehicle, xmin = xmin, xmax = xmax, ymin = ymin,
                   ymax = ymax)) + 
  geom_rect() +
  geom_text(data = subset(flyers, geom == "geom_text"),
            aes(x = (xmin + xmax) / 2, y = (ymin + ymax) / 2)) +
  geom_fit_text(data = subset(flyers, geom == "geom_fit_text")) +
  facet_wrap( ~ geom) +
  labs(x = "", y = "")
```

There are three different ways to define the box in which you want the text to
be placed:

1. On a continuous axis, you can use `xmin`/`xmax` and/or `ymin`/`ymax`
   aesthetics.
2. Alternatively on a continuous axis, you can define the horizontal and/or
   vertical centre of the box with `x` and/or `y` respectively, and fix the
   width and/or height of the box with the `width` and/or `height` arguments.
   The values for `width` and `height` should be `grid::unit()` objects; if
   not, they will be assumed to use the native axis scale.
3. On a discrete (categorical) axis, `geom_fit_text()` will automatically
   figure out the appropriate width or height. You can override this with the
   `width` or `height` arguments if you want.

You can specify where in the box to place the text with the `place` argument,
and a minimum point size for the text with the `min.size` argument. (Any text
that would need to be smaller than `min.size` to fit the box will be hidden.)

```{r geom_fit_text_2}
ggplot(flyers, aes(label = vehicle, xmin = xmin, xmax = xmax, ymin = ymin,
                   ymax = ymax)) + 
  geom_rect() +
  geom_text(data = subset(flyers, geom == "geom_text"),
            aes(x = (xmin + xmax) / 2, y = (ymin + ymax) / 2)) +
  geom_fit_text(data = subset(flyers, geom == "geom_fit_text"),
                place = "top", min.size = 6) +
  facet_wrap( ~ geom) +
  labs(x = "", y = "")
```

Text can be placed in any corner (‘topleft’, ‘topright’, …) or at the midpoint
of any side (‘bottom’, ‘left’, …), as well as the default ‘centre’.

# Growing text

With the `grow = TRUE` argument, text will be made to fill the box completely,
whether that requires growing or shrinking it.

```{r geom_fit_text_3}
ggplot(flyers, aes(label = vehicle, xmin = xmin, xmax = xmax, ymin = ymin, 
                   ymax = ymax)) +
  geom_rect() +
  geom_text(data = subset(flyers, geom == "geom_text"),
            aes(x = (xmin + xmax) / 2, y = (ymin + ymax) / 2)) +
  geom_fit_text(data = subset(flyers, geom == "geom_fit_text"), grow = TRUE) +
  facet_wrap( ~ geom, ncol = 1) +
  labs(x = "", y = "")
```

## Reflowing text

With the `reflow = TRUE` argument, text will be reflowed (wrapped) as needed to
fit the box. Reflowing is preferred to shrinking; that is, if the text can be
made to fit by reflowing it without shrinking it, it will be reflowed only.

```{r reflow}
poem <- data.frame(
  text = rep(
    "Whose words these are I think I know.\nHe would prefer that they reflow",
    3
  ),
  xmin = rep(10, 3),
  xmax = rep(90, 3),
  ymin = rep(10, 3),
  ymax = rep(90, 3),
  fit = c("geom_text", "without reflow", "with reflow")
)

ggplot(poem, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                 label = text)) +
  geom_rect() +
  geom_text(
    data = subset(poem, fit == "geom_text"),
    aes(x = (xmin + xmax) / 2, y = (ymin + ymax) / 2)
  ) +
  geom_fit_text(data = subset(poem, fit == "without reflow"), min.size = 0) +
  geom_fit_text(data = subset(poem, fit == "with reflow"), reflow = TRUE,
                min.size = 0) +
  lims(x = c(0, 100), y = c(0, 100)) +
  labs(x = "", y = "") +
  facet_wrap(~ fit)
```

Note that existing line breaks in the text are respected.

With both `reflow = TRUE` and `grow = TRUE`, the text will be reflowed to a
form that best matches the aspect ratio of the box, then made to fill the box
whether that requires shrinking it or growing it.

```{r reflow_and_grow}
film <- data.frame(
  text = rep("duck soup", 3),
  xmin = rep(30, 3),
  xmax = rep(70, 3),
  ymin = rep(0, 3),
  ymax = rep(100, 3),
  fit = c("geom_text", "grow without reflow", "grow with reflow")
)

ggplot(film, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                 label = text)) +
  geom_rect() +
  geom_text(
    data = subset(film, fit == "geom_text"),
    aes(x = (xmin + xmax) / 2, y = (ymin + ymax) / 2)
  ) +
  geom_fit_text(data = subset(film, fit == "grow without reflow"), grow = TRUE) +
  geom_fit_text(
    data = subset(film, fit == "grow with reflow"),
    grow = TRUE,
    reflow = TRUE
  ) +
  lims(x = c(0, 100), y = c(0, 100)) +
  labs(x = "", y = "") +
  facet_wrap(~ fit, ncol = 1)
```

## Example: a heatmap

```{r heatmap, fig.width = 8}
tournament <- data.frame(
  teamA = rep(c("Radon Canyon Raiders", "Desert Bluffs Destroyers",
                "Old Town Orphans"), 3),
  teamB = c(rep("Coyote Corners Carolers", 3),
            rep("Sand Wastes Streetcleaners", 3),
            rep("Dog park Devils", 3)),
  venue = sample(c("Night Vale Stadium", "Big Rico's Sandlot With The Lot",
                   "Elementary School Playground",
                   "Night Vale Harbor and Waterfront Recreation Area"),
                 9, replace = TRUE),
  game_time_mins = sample(999, 9)
  )

ggplot(tournament, aes(x = teamA, y = teamB, fill = game_time_mins,
                       label = venue)) +
  geom_tile() +
  geom_fit_text(reflow = TRUE, grow = TRUE, colour = "white")
```
